# Noncompliance in Recycling Experiment 

```{problem_description}
Suppose that you want to conduct a study of recycling behavior. A number of undergraduate students are hired to walk door to door and provide information about the benefits of recycling to people in the treatment group. Here are some facts about how the experiment was actually carried out.  

- 1,500 households are assigned to the treatment group. 
- The undergrads tell you that they successfully managed to contact 700 households. 
- The control group had 3,000 households (not contacted by any undergraduate students). 
- The subsequent recycling rates (i.e. the outcome variable) are computed and you find that 500 households in the treatment group recycled. In the control group, 600 households recycled.
```

## Intent to treat effect
What is the ITT? Do the work to compute it, and store it into the object `recycling_itt`.  Provide a short narrative using inline R code, such as `r inline_reference`.

```{r recycling itt, include=TRUE} 
'ITT = E[Y | Z = 1] − E[ Y | Z = 0]' 
group_a = (600/1500)
#group_a
group_b = (800/1500)
#group_b
group_c = (3000/3000)
#group_c
Y_t_a = (500/600)
#Y_t_a
Y_t_c = (600/3000)
#Y_t_c
#I do not believe we can use the control group percentage for those that were assigned to treatment but received a dose of control.
#recycling_itt_b <- (Y_t_a*group_a)+(Y_t_c*group_b) - (Y_t_c)
#recycling_itt_b
recycling_itt <- (Y_t_a*group_a)+(0*group_b) - (Y_t_c)
recycling_itt
```

**Answer:** The ITT is 0.133, which tells us that the effect if receiving information about the benefits of recycling results in a 13.3% increase in recycling as compared to not receiving such information. This tells us the treatment effect for treatment/control assignments, regardless of compliance.

## Compliers average causal effect
What is the CACE? Do the work to compute it, and store it into the object `recycling_cace`.  Provide a short narrative using inline R code.

```{r recycling cace, include=TRUE} 
ITT_D = (700/1500)
recycling_cace <- recycling_itt/ITT_D
recycling_cace
```

**Answer:** The CACE is 0.286, which tells us that, for compliers, receiving information about recycling resulted in a 51.4% increase in recycling as compared to not receiving the information.

```{problem_description}
There appear to be some inconsistencies regarding how the undergraduates actually carried out the instructions they were given. 

- One of the students, Mike, tells you that they actually lied about the the number of contacted treatment households and that the true number was 500. 
- Another student, Andy, tells you that the true number was actually 600.
```

## Mike's CACE
What is the CACE if Mike is correct?  Provide a short narrative using inline R code.

```{r cace mike, include=TRUE} 
'ITT = E[Y | Z = 1] − E[ Y | Z = 0]' 
group_a = (500/1500)
group_b = (1000/1500)
group_c = (3000/3000)
Y_t_a = (500/500)
Y_t_c = (600/3000)

recycling_itt <- (Y_t_a*group_a)+(0*group_b) - (Y_t_c)
ITT_D = (500/1500)
cace_mike <- recycling_itt/ITT_D
cace_mike
```

**Answer:** The CACE is 0.4, which tells us that as a result of receiving information on recycling people are 40% more likely to recycle as compared to not receiving this information. This is a ~1.2% increase in effect size as compared to the original numbers.

## Andy's CACE 
What is the CACE if Andy is correct? Provide a short narrative using inline R code.

```{r cace andy, include=TRUE}
'ITT = E[Y | Z = 1] − E[ Y | Z = 0]' 
group_a = (600/1500)
group_b = (900/1500)
group_c = (3000/3000)
Y_t_a = (500/600)
Y_t_c = (600/3000)

recycling_itt <- (Y_t_a*group_a)+(0*group_b) - (Y_t_c)
ITT_D = (600/1500)
cace_andy <- recycling_itt/ITT_D
cace_andy
```

**Answer:** The CACE is 0.333, which tells us that as a result of receiving information on recycling people are 40% more likely to recycle as compared to not receiving this information. This is a ~0.7% reduction in effect size as compared to the Mike's numbers but still ~0.5% increase from the original numbers.

```{problem_description}
For the rest of this question, suppose that **in fact** Mike was telling the truth.  
```

## Effect of false reporting
What was the impact of the undergraduates’s false reporting on our estimates of the treatment’s effectiveness?

**Answer:** The impact was that our CACE estimate was below what the true estimate may be. If we are to believe either mike or andy, then we woul have underestimated the true estimate by ~0.5%-1.2%. 

## Effect of false reporting... on what quantity?
Does your answer change depending on whether you choose to focus on the ITT or the CACE?

**Answer:** The ITT remains the same but CACE changes. Given that we do not have any information on non-compliers for treatment, the ITT is not changed.



